### 2021-08-24 text/article_vedrovice.Rmd
update text before submission

---
title           : "Spatial Patterns and Grave Goods Differences at the Cemetery of Vedrovice (Czech Republic): A Resampling Approach to Identity Markers in the Early Neolithic"
author          : "Petr Pajdla"
date            : "`r format(Sys.time(), '%d %B, %Y')`"
abstract        :
  "The spatial pattern and grave good associations at the cemetery of Vedrovice - Široká u Lesa are explored using computationally intensive formal methods. Resampling approaches to both intra-site spatial arrangement and the distribution of grave goods are explored in relation to layers of identity represented by different buried body attributes – sex, age, geographical origin and their combinations. We found that polished stone adzes, grinding tools, ceramic bowls and spondylus beads occur among non-randomly distributed grave goods. As a result of the investigation of differences among the non-randomly distributed artifacts and attributes of burials, we conclude that the most significant are dichotomies between genders, accentuated by both origin and age. Spatial analysis shows clear trends towards clustering between burials, and the emerging clusters are further explored for similarities in burial attributes."
output:
    bookdown::word_document2:
        latex: pdflatex
        toc: false
        fig_caption: true
        number_sections: false
        keep_tex: false
bibliography    : "bib_vedrovice.bib"
csl             : "apa.csl"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
---

@@

# Introduction

@@

<!-- view of gender in the Neolithic, archaeological context -->
@robb2017 [p. 128] noted a lack of debate in archaeology on gender for the
European Neolithic compared to other periods.
The authors present gender in the Neolithic as being
of a very different nature than gender in later prehistory.
It is argued that gender is not clearly expressed and
biological differences between sexes are not transformed into binary cultural
categories [@robb2017, p. 140--142].
Considering the Neolithic burial evidence across Europe, authors find that it
"shows very little gender distinction; when it does, it is rarely a clear-cut
binary categorization but a matter of degree and overlap, sometimes polythetically
combined with factors such as age and localness" [@robb2017, p. 138].
In response, @bickle2020 [p. 14] argues that "biological sex may not have defined gender
during the LBK" although the "sexed bodies did matter".

@@

# Material and methods

@@

Another category of variables is derived from retrieved human skeletal remains.
These include biological sex, age simplified into age groups and an indicator
variable as to whether the strontium signal in the remains is of local origin.
The variable representing the recorded sex is simplified into the following
categories: female-sexed body (abbreviated *F*), male-sexed body (*M*),
non-adult (*n. a.*) and undetermined sex (*ind.*).
The age of the body is also simplified into three categories: juveniles (*juv.*)
comprised of the groups *Infans* and *Juvenis*, adults (*ad.*, all bodies aged as *Adultus*)
and mature bodies (*mat.*, anthropological categories *Adultus-Maturus*, *Maturus* and
*Senilis*) and bodies of an undetermined age (*ind.*).
The last category is the indicator for the origin of the buried person.
The local strontium range is defined as two standard deviations from the mean of Sr ranges of
bodies with an age category determined as *Infans*.
Bodies falling within this range are labeled as local, outside the range as non-local;
where data on Sr values is missing, origin is not determined (*ind.*).

@@

## Randomness of artifact combinations

@@

<!-- spatial analysis -->
## Spatial analysis

@@

To further explore the neighborhood structure, we construct neighborhood graphs.
We use Delaunay triangulation and its subgraph, a Gabriel graph [@gabriel1969]
with burial locations as nodes (Fig. \@ref(fig:plan-graphs)).
The Delaunay triangulation is regarded as the maximum number of neighbors per individual case
scenario, while the Gabriel graph is the opposite.
For each node, the mean number of neighbors of a given category is determined and recorded.
To check how common the observed value is, a permutation test is performed
where the categories are randomly assigned while preserving the overall
proportion of given categories and the neighborhood structure determined by the
graph.
The observed values of neighbor categories are compared to distributions
derived from 999 permutations of the simulated data.

@@

<!-- The size of the radius is based on the results of nearest neighbors analysis, where
at the given radius there is strong evidence of a clustering process (Fig. \@ref(fig:pp-funs)).
The randomness is again tested using a permutation test similar to the one
described for the neighborhood derived from the graphs.-->

<!-- software packages -->
Data is analyzed in an R environment for statistical computing and graphics
[@rcoreteam2021] using functions of several different packages; 
the most frequently used are packages from the *tidyverse* family [@wickham2019a].
Packages *sf* [@pebesma2018], *spatstat* [@baddeley2015], *spdep* [@bivand2018] and 
*concaveman* [@gombin2020] are used to perform various spatial analysis tasks; 
*igraph* [@csardi2006a] to construct and analyze networks;
*vegan* [@oksanen2020] and *randomizr* [@coppock2019] in the randomization experiments.

@@

## Artifact combinations

By counting the co-occurrences of analyzed variables on simulated data matrices and comparing
the distributions of the statistics proposed by @manly1996 with the values calculated for
the original data set, we can establish whether the overall pattern of artifact
distribution in the graves is non-random and further identify artifacts that deviate
from the random pattern.
The overall deviation from expected co-occurrences, i.e., the *S* statistic
[@manly1996, 475] is significantly large at the 0.05 alpha level.
This suggests that there is strong evidence that the overall distribution of
artifacts in different graves is not random. <!--(see Fig. \@ref(fig:sstat))-->

Examining distributions of individual artifacts, we find that at the 0.05
significance level only adzes, grinding tools, ceramic bowls and spondylus beads
are distributed non-randomly.
Close to the 0.05 alpha level are distributions of bone tools and ceramic bottles
(Fig. \@ref(fig:cooc-vstat)).
Among the grave goods found to be randomly distributed in the graves
across the burial ground are ocher pigment, lithics, pebbles, globular pots and
special types of pottery vessel shapes, spondylus pendants and buckles.

The relationships between artifact types frequently occurring and missing together
in graves are visualized in Fig. \@ref(fig:cooc-nets) with
clusters established using a fast-greedy algorithm [@newman2004] and the strength of the
given association expressed by the thickness of the edge connecting the nodes.
Artifacts most frequently present with other artifacts are grinding tools and
ceramic bowls with five links and adzes, non-local lithic tools, ocher pigment and
spondylus beads with four links.
The association is strongest between spondylus beads and (a) ceramic bottles and
(b) pigment; and adzes and (a) non-local lithics and (b) bone tools.
Artifacts frequently absent with other artifacts in the graves are  
globular pots, local lithic tools, ceramic bowls and spondylus buckles
with seven to four links respectively.
The strongest antagonist relationships are between globular pots and (a) bowls and
(b) adzes.

@@

## Spatial patterning

<!-- The quadrat test shows evidence for both random and clustered processes taking place with p values
under the 0.05 alpha level but the $\chi^2$ approximation is inaccurate due to small
expected counts (for the details see supplementary information).
Inconclusive result of the quadrat test is not surprising because as was noted
earlier, we do not expect *first-order* properties to play major roles in
placing of the individual burials. -->
The results of functions based on nearest neighbors are reported in Fig. \@ref(fig:pp-funs).
Both *G* and  *F* functions slightly deviate from the curve expected
under an independent random process,
but neither crosses the boundaries of the envelope.
Thus, the cumulative nearest-neighbor distance (*G*) as well as the
empty-space function (*F*) are well approximated by
the random process and do not show evidence for spatial clustering or repellent functions
taking place.
On the contrary,
the *K* and *T* functions show clear deviations from the expected curve and both exceed
the boundaries of the given envelopes at distances around or smaller than 5 meters.
<!-- in case of *K* and slightly more in case of *T* function.-->
This means that the number of nearest neighbors at this and larger distances is
significantly higher than we would expect under complete spatial randomness.
The overall point pattern thus shows a clear signs of clustering process taking place.

@@

The results of a percolation analysis reveal a more detailed view of the clustering process
taking place; see Fig. \@ref(fig:perc-distance).
The distances at which the relative stability of clusters emerges are labeled.
The curve reveals a steady growth up until a
radius of 5 meters and then a large skip towards a single cluster
between 5.5 and 6.5 meters.
This point of cluster growth is revealed by the point pattern analysis as well.
The clusters at chosen distances are shown in Fig. \@ref(fig:perc-plan).

## Neighborhood structure

The results of the multitype *K* function show deviations from random processes for
most of the observed categories in Fig. \@ref(fig:pp-crossk).
The distances at which these deviations start occurring are in most cases rather large,
well over 10 meters and, considering the fact that at the 10.4 meters radius
around individual points the percolation analysis proved the existence
of a single large cluster encompassing almost all of the points, these
observations correspond to the deviations observed for the single *K* function.
The only notable deviation we consider noteworthy is in the case of age
category in pairs of juveniles and in the sex category between pairs of non-adults.
Here the deviation from random processes is observed at smaller distances than
in the case of other variables at distances around 5 meters.
This proves a tighter bound cluster of juveniles in the southeast part of the cemetery.

@@

In the case of neighborhood structure derived from Delaunay triangulation (maximum
number of neighbors scenario) and
Gabriel graph (minimum number of neighbors scenario, Fig. \@ref(fig:plan-graphs)),
most of the observed relationships are not significantly different from those
derived by randomization experiment.
In the Delaunay triangulation the significant relationships are observed in two cases
between (1) pairs of non-adults and (2) female-sexed bodies and non-adults.
The first observation can again be connected with a higher concentration of
non-adults in the southeast space of the burial ground.
In regard to the sex category, the Gabriel graph shows an association
between female-sexed and male-sexed bodies.
In only one case are the observations from both of the graphs significant, and
that is among pairs of bodies in the mature age category.

@@

Clusters formed at different percolation distances
show the following deviations from the expected number of
occurrences of given categories at different scales.
Regarding the determination of the sex of the bodies, at the threshold of 3.9 meters
we observe that in the largest cluster (no. 3, Fig. \@ref(fig:perc-plan))
there are more non-adults and less male-sexed bodies and in the second largest
cluster (no. 10) there are slightly more male-sexed bodies than would
be expected under a random process.
The over-representation of non-adult bodies is repeated in the following
threshold distance of 4.6 meters in the largest cluster (no. 3).
Last but not least, in cluster no. 8 (north of the cemetery) at the distance
of 5.5 meters, female-sexed bodies are more frequent than we would expect.
The age categories correspond to the previous observations where
the over-representation of juvenile bodies at the 3.9-meter threshold in
cluster no. 3 is most evident.
Exploring the origin of the bodies based on the strontium isotope signature, we
find that the cluster in the southwest part at threshold distances of 3.9 and 5.5 meters
(no. 10 and 5 respectively) consists of less non-local bodies than would be expected.
Finally, at a radius of 2.8 meters there are more local bodies in the central
cluster of the cemetery (no. 8).

@@

<!-- Methods of neighborhood relationship construction examined include
distance based approaches where all the neighbors that fall into a given buffer are considered,
the other two are based on neighborhood graphs, the Delaunay triangulation as a
maximum number of neighbors scenario and a Gabriel graph as a minimal scenario. -->
<!--
For relationships (see Fig. \@ref(fig:nb-sex) and Tab. \@ref(tab:nb-tab-sex))
of male-sexed bodies, none of the methods showed any significant
differences between expected and observed number of neighbors of a given sex.
The relationship with female-sexed bodies is rather extreme for all of the methods employed,
but never significant ($p > 0.07$).
In case of bodies unsexed due to early age (non-adults), the neighborhood based
on buffer zones show a significant relationship with other non-adults ($p \approx 0.04$).
Finally burials with female-sexed bodies show a significantly higher number of
burials with male-sexed bodies in both the methods based on neighborhood networks
(Delaunay triangulation $p \approx 0.02$ and Gabriel graph $p \approx 0.03$).
Two further relationships showed non-random deviations.
In case of Delaunay triangulation for burials with bodies with indeterminate sex with
non-adults ($p \approx 0.02$) and in case of Gabriel graph between non-adults and
indeterminate category $p \approx 0.02$.
-->

@@

# Conclusions and discussion

@@

@@

```{r fig-widths}
fig_wid <- c(".3\\textwidth",
             ".6\\textwidth",
             ".8\\textwidth",
             ".98\\textwidth")
```

@@

```{r label-pp-funs}
label_pp_funs <- paste(
  "Functions *G*, *F*, *K* and *T*.",
  "*Dashed line* - theoretical curve under complete spatial randomness;
  *solid line* - empirical curve; *gray area* - envelope")
```
```{r pp-funs, fig.cap=label_pp_funs, out.width=fig_wid[4]}
knitr::include_graphics(here("plots", "pp_funs.pdf"))
```

>>>>

---
title           : "Spatial Patterns and Grave Goods Differences at the Cemetery of Vedrovice (Czech Republic): A Resampling Approach to Identity Markers in the Early Neolithic"
author          : ""
date            : "`r format(Sys.time(), '%d %B, %Y')`"
abstract        :
  "The spatial pattern and grave good associations at the cemetery of Vedrovice - Široká u Lesa are explored using computationally intensive formal methods. Resampling approaches to both intra-site spatial arrangement and the distribution of grave goods are explored in relation to layers of identity represented by different buried body attributes – sex, age, geographical origin and their combinations. We found that polished stone adzes, grinding tools, ceramic bowls and spondylus beads occur among non-randomly distributed grave goods. As a result of the investigation of differences among the non-randomly distributed artifacts and attributes of burials, we conclude that the most significant are dichotomies between genders, accentuated by both origin and age. Spatial analysis shows clear trends towards clustering between burials, and the emerging clusters are further explored for similarities in burial attributes."
output:
    bookdown::word_document2:
        latex: pdflatex
        toc: false
        fig_caption: true
        number_sections: false
        global_numbering: false
        keep_tex: false
bibliography    : "bib_vedrovice.bib"
csl             : "journal-of-computer-applications-in-archaeology.csl"
---

@@

# Introduction {-}

@@

<!-- view of gender in the Neolithic, archaeological context -->
@robb2017 [p. 128] noted a lack of debate in archaeology on gender for the
European Neolithic compared to other periods.
The authors present gender in the Neolithic as being
of a very different nature than gender in later prehistory.
It is argued that gender is not clearly expressed and
biological differences between sexes are not transformed into binary cultural
categories [@robb2017, p. 140--142].
Considering the Neolithic burial evidence across Europe, authors find that it
'shows very little gender distinction; when it does, it is rarely a clear-cut
binary categorization but a matter of degree and overlap, sometimes polythetically
combined with factors such as age and localness' [@robb2017, p. 138].
In response, @bickle2020 [p. 14] argues that 'biological sex may not have defined gender
during the LBK' although the 'sexed bodies did matter'.

@@

# Material and Methods

@@

Another category of variables is derived from retrieved human skeletal remains.
These include biological sex, age simplified into age groups and an indicator
variable as to whether the strontium signal in the remains is of local origin.
The variable representing the recorded sex is simplified into the following
categories: female-sexed body (abbreviated *F*), male-sexed body (*M*),
non-adult (*n. a.*) and undetermined sex (*ind.*); see Figure \@ref(fig:plan-sex).
The age of the body is also simplified into three categories: juveniles (*juv.*)
comprised of the groups *Infans* and *Juvenis*, adults (*ad.*, all bodies aged as *Adultus*)
and mature bodies (*mat.*, anthropological categories *Adultus-Maturus*, *Maturus* and
*Senilis*) and bodies of an undetermined age (*ind.*); see Figure \@ref(fig:plan-age).
The last category is the indicator for the origin of the buried person.
The local strontium range is defined as two standard deviations from the mean of Sr ranges of
bodies with an age category determined as *Infans*.
Bodies falling within this range are labeled as local, outside the range as non-local;
where data on Sr values is missing, origin is not determined (*ind.*); see Figure \@ref(fig:plan-orig).

<!--
[Figure_04]

[Figure_05]

[Figure_06]
-->

@@

## Randomness of Artifact Combinations

@@

<!-- spatial analysis -->
## Spatial Analysis

@@

To further explore the neighborhood structure, we construct neighborhood graphs.
We use Delaunay triangulation and its subgraph, a Gabriel graph [@gabriel1969]
with burial locations as nodes (Figure \@ref(fig:plan-graphs)).
The Delaunay triangulation is regarded as the maximum number of neighbors per individual case
scenario, while the Gabriel graph is the opposite.
For each node, the mean number of neighbors of a given category is determined and recorded.
To check how common the observed value is, a permutation test is performed
where the categories are randomly assigned while preserving the overall
proportion of given categories and the neighborhood structure determined by the
graph.
The observed values of neighbor categories are compared to distributions
derived from 999 permutations of the simulated data.

<!--[Figure_09]-->

@@

<!-- The size of the radius is based on the results of nearest neighbors analysis, where
at the given radius there is strong evidence of a clustering process (Figure \@ref(fig:pp-funs)).
The randomness is again tested using a permutation test similar to the one
described for the neighborhood derived from the graphs.-->

<!-- software packages -->
Data is analyzed in an R environment for statistical computing and graphics
[@rcoreteam2021] using functions of several different packages; 
the most frequently used are packages from the *tidyverse* family [@wickham2019].
Packages *sf* [@pebesma2018], *spatstat* [@baddeley2015], *spdep* [@bivand2018] and 
*concaveman* [@gombin2020] are used to perform various spatial analysis tasks; 
*igraph* [@csardi2006a] to construct and analyze networks;
*vegan* [@oksanen2020] and *randomizr* [@coppock2019] in the randomization experiments.

@@

## Artifact Combinations

By counting the co-occurrences of analyzed variables on simulated data matrices and comparing
the distributions of the statistics proposed by @manly1996 with the values calculated for
the original data set, we can establish whether the overall pattern of artifact
distribution in the graves is non-random and further identify artifacts that deviate
from the random pattern.
The overall deviation from expected co-occurrences, i.e., the *S* statistic
[@manly1996, 475] is significantly large at the 0.05 alpha level.
This suggests that there is strong evidence that the overall distribution of
artifacts in different graves is not random (see Figure \@ref(fig:sstat)).

<!--[Figure_01]-->

Examining distributions of individual artifacts, we find that at the 0.05
significance level only adzes, grinding tools, ceramic bowls and spondylus beads
are distributed non-randomly.
Close to the 0.05 alpha level are distributions of bone tools and ceramic bottles
(Figure \@ref(fig:cooc-vstat)).
Among the grave goods found to be randomly distributed in the graves
across the burial ground are ocher pigment, lithics, pebbles, globular pots and
special types of pottery vessel shapes, spondylus pendants and buckles.

<!--[Figure_02]-->

The relationships between artifact types frequently occurring and missing together
in graves are visualized in Figure \@ref(fig:cooc-nets) with
clusters established using a fast-greedy algorithm [@newman2004] and the strength of the
given association expressed by the thickness of the edge connecting the nodes.
Artifacts most frequently present with other artifacts are grinding tools and
ceramic bowls with five links and adzes, non-local lithic tools, ocher pigment and
spondylus beads with four links.
The association is strongest between spondylus beads and (a) ceramic bottles and
(b) pigment; and adzes and (a) non-local lithics and (b) bone tools.
Artifacts frequently absent with other artifacts in the graves are  
globular pots, local lithic tools, ceramic bowls and spondylus buckles
with seven to four links respectively.
The strongest antagonist relationships are between globular pots and (a) bowls and
(b) adzes.

<!--[Figure_03]-->

@@

## Spatial Patterning

<!-- The quadrat test shows evidence for both random and clustered processes taking place with p values
under the 0.05 alpha level but the $\chi^2$ approximation is inaccurate due to small
expected counts (for the details see supplementary information).
Inconclusive result of the quadrat test is not surprising because as was noted
earlier, we do not expect *first-order* properties to play major roles in
placing of the individual burials. -->
The results of functions based on nearest neighbors are reported in Figure \@ref(fig:pp-funs).
Both *G* and  *F* functions slightly deviate from the curve expected
under an independent random process,
but neither crosses the boundaries of the envelope.
Thus, the cumulative nearest-neighbor distance (*G*) as well as the
empty-space function (*F*) are well approximated by
the random process and do not show evidence for spatial clustering or repellent functions
taking place.
On the contrary,
the *K* and *T* functions show clear deviations from the expected curve and both exceed
the boundaries of the given envelopes at distances around or smaller than 5 meters.
<!-- in case of *K* and slightly more in case of *T* function.-->
This means that the number of nearest neighbors at this and larger distances is
significantly higher than we would expect under complete spatial randomness.
The overall point pattern thus shows a clear signs of clustering process taking place.

<!--[Figure_07]-->

@@

<!--[Figure_10]-->

The results of a percolation analysis reveal a more detailed view of the clustering process
taking place; see Figure \@ref(fig:perc-distance).
The distances at which the relative stability of clusters emerges are labeled.
The curve reveals a steady growth up until a
radius of 5 meters and then a large skip towards a single cluster
between 5.5 and 6.5 meters.
This point of cluster growth is revealed by the point pattern analysis as well.
The clusters at chosen distances are shown in Figure \@ref(fig:perc-plan).

<!--[Figure_11]-->

## Neighborhood Structure

The results of the multitype *K* function show deviations from random processes for
most of the observed categories in Figure \@ref(fig:pp-crossk).
The distances at which these deviations start occurring are in most cases rather large,
well over 10 meters and, considering the fact that at the 10.4 meters radius
around individual points the percolation analysis proved the existence
of a single large cluster encompassing almost all of the points, these
observations correspond to the deviations observed for the single *K* function.
The only notable deviation we consider noteworthy is in the case of age
category in pairs of juveniles and in the sex category between pairs of non-adults.
Here the deviation from random processes is observed at smaller distances than
in the case of other variables at distances around 5 meters.
This proves a tighter bound cluster of juveniles in the southeast part of the cemetery.

<!--[Figure_08]-->

@@

In the case of neighborhood structure derived from Delaunay triangulation (maximum
number of neighbors scenario) and
Gabriel graph (minimum number of neighbors scenario, Figure \@ref(fig:plan-graphs)),
most of the observed relationships are not significantly different from those
derived by randomization experiment.
In the Delaunay triangulation the significant relationships are observed in two cases
between (1) pairs of non-adults and (2) female-sexed bodies and non-adults.
The first observation can again be connected with a higher concentration of
non-adults in the southeast space of the burial ground.
In regard to the sex category, the Gabriel graph shows an association
between female-sexed and male-sexed bodies.
In only one case are the observations from both of the graphs significant, and
that is among pairs of bodies in the mature age category.

@@

Clusters formed at different percolation distances
show the following deviations from the expected number of
occurrences of given categories at different scales.
Regarding the determination of the sex of the bodies, at the threshold of 3.9 meters
we observe that in the largest cluster (no. 3, Figure \@ref(fig:perc-plan))
there are more non-adults and less male-sexed bodies and in the second largest
cluster (no. 10) there are slightly more male-sexed bodies than would
be expected under a random process.
The over-representation of non-adult bodies is repeated in the following
threshold distance of 4.6 meters in the largest cluster (no. 3).
Last but not least, in cluster no. 8 (north of the cemetery) at the distance
of 5.5 meters, female-sexed bodies are more frequent than we would expect.
The age categories correspond to the previous observations where
the over-representation of juvenile bodies at the 3.9-meter threshold in
cluster no. 3 is most evident.
Exploring the origin of the bodies based on the strontium isotope signature, we
find that the cluster in the southwest part at threshold distances of 3.9 and 5.5 meters
(no. 10 and 5 respectively) consists of less non-local bodies than would be expected.
Finally, at a radius of 2.8 meters there are more local bodies in the central
cluster of the cemetery (no. 8).

@@

<!-- Methods of neighborhood relationship construction examined include
distance based approaches where all the neighbors that fall into a given buffer are considered,
the other two are based on neighborhood graphs, the Delaunay triangulation as a
maximum number of neighbors scenario and a Gabriel graph as a minimal scenario. -->
<!--
For relationships (see Figure ref(fig:nb-sex) and Tab. ref(tab:nb-tab-sex))
of male-sexed bodies, none of the methods showed any significant
differences between expected and observed number of neighbors of a given sex.
The relationship with female-sexed bodies is rather extreme for all of the methods employed,
but never significant ($p > 0.07$).
In case of bodies unsexed due to early age (non-adults), the neighborhood based
on buffer zones show a significant relationship with other non-adults ($p \approx 0.04$).
Finally burials with female-sexed bodies show a significantly higher number of
burials with male-sexed bodies in both the methods based on neighborhood networks
(Delaunay triangulation $p \approx 0.02$ and Gabriel graph $p \approx 0.03$).
Two further relationships showed non-random deviations.
In case of Delaunay triangulation for burials with bodies with indeterminate sex with
non-adults ($p \approx 0.02$) and in case of Gabriel graph between non-adults and
indeterminate category $p \approx 0.02$.
-->

@@

# Conclusions and Discussion {-}

@@

<!-- Acknowledgement / Reproducibility -->

# Reproducibility {-}

To enable re-use and reproducibility, a compendium containing data and code accompanying this paper is released at Zenodo data repository under DOI [10.5281/zenodo.4899652)](https://doi.org/10.5281/zenodo.4899652).

# Acknowledgements {-}

The research described in this paper was accomplished with support from the project ‘Lifestyle as an unintentional identity in the Neolithic’ (Project 19-16304S), financed by the Czech Science Foundation.

@@

```{r fig-widths}
fig_wid <- c(".3\\textwidth",
             ".6\\textwidth",
             ".8\\textwidth",
             ".9\\textwidth")
```

@@

```{r label-pp-funs}
label_pp_funs <- paste("Functions G, F, K and T. Dashed line – theoretical curve under complete spatial randomness; solid line – empirical curve; gray area – envelope")
```
```{r pp-funs, fig.cap=label_pp_funs, out.width=fig_wid[4]}
knitr::include_graphics(here("plots", "pp_funs.pdf"))
```
